### 2013-04-01 doc/icml2013/rebuttal.txt

Thank you for your constructive comments and suggestions; we will try to
respond to your questions below.

# Major Comments:

## Reviewer 30 (Regarding Lemma 3):
The $O(1/\sqrt{n})$ convergence follows from applying Hoeffding's
inequality on each term of the random vectors $\eta_p(x) x$, which have
zero mean when taking expectation over $h$ and $\epsilon$. For example,
in $\eta_2(x) x$, we have three terms (Equation 3), which each have
expectation zero by independence of $x$, $h$ and $\epsilon$.

This does not require that the $x$ have zero mean, and note that we do
not take expectations over $x$ at any point in our analysis.

## Reviewer 4 (Regarding analysis with whitening):
Your point is well noted; results from standard techniques (Hsu
& Kakade, 2013) show that the whitening step introduces an additional
factor of $O(\sigma_{min}(M_2)^{-1.5} \|M_2\|_{op}^{0.5} \|M_3\|_{op})$.
We will include this as a lemma in the final version as we can not
upload an updated version in the rebuttal phase.

## Reviewer 30 (Regarding convergence rate in experiments):

One of the obstacles in applying the bounds we derived is that it
requires the error of the empirical moment estimates to be less than
$\sigma_{\min}(M_2)/2$, a condition that did not hold true for our
experiments. For the data in Figure 3, this condition required the
moment errors to be less than 0.15, while the error in estimates we
recovered were at least 0.9. We have also identified that one major
reason for this was that the proximal gradient algorithm we used did not
converge in the specified number of iterations (i.e. 500).

# Minor Comments:

## Reviewer 30 (Regarding Identifiability):
Identifiability is only a problem when the input features are not
algebraically independent, i.e. each coordinate can not be expressed as
a deterministic analytic function of the remaining coordinates. Thus, it
is identifiable when the coordinates are generated from a random
Gaussian (or any other distribution).

## Reviewer 4 (Regarding Lemma 3):
We do not need to know $h$; in fact we are trying to bound the
difference between mean, $M_1$ which we recover by regression, and the
actual value of $\beta_h$ at each data point; it is useful to think of
it as bounding the variance of the zero-mean term $M_1 - \beta_h$.

## Reviewer 30 (Regarding operator norm for vectors):
This was a typo, and we meant the 2-norm.

## Reviewer 4 (Regarding references):
Thank you, we will include them in the final version.

## Reviewer 4 (Regarding $\sigma$):
We used $\sigma = 0.1$ for the results we presented, but found that they
were not very sensitive to the noise level.

## Reviewer 4 (Regarding Regularization):
We set the regularization parameters by grid search for $n = 10^5$ and
used the $1/\sqrt{n}$ rate to extrapolate for other $n$.

## Reviewer 7 (Regarding Generality of Experiments):
We found that robust initialization happened for almost all inputs. The
cases where we did not get robust initialization occurred when we did
not have good enough moment estimates and thus poor parameter estimates.
Furthermore, this problem went away when we used more samples.

## Reviewer 7 (Regarding Plot of Convergence):
We have compared the convergence of spectral experts and EM to the exact
parameters for increasing values of $n$ in Figure 3 in terms of the parameter
error.

### 2013-04-23 doc/icml2013/algo.tex
Couple of corrections to the math in the theory and introduction section. Included a line about mode-k tensor unfoldings. Added details about $\sigma^2$ to the experiments

%%% first moment
To warm up, let us consider linear regression
on the response $y$ given $x$.
From the model definition, we have $y = \beta_h^\top x + \epsilon$ where
$\epsilon \sim \normal{0}{\sigma^2}$, and $h$ is a random quantity
independent of $x$.
We can average over this randomness by defining
the average regression coefficients
$M_1 \eqdef \sum_{h=1}^k \pi_h \beta_h$.
Now we can express $y$ as a linear function of $x$ with coefficients $M_1$
plus some noise $\eta_1(x)$:
\begin{align}
  y &= \innerp{M_1}{x} +
  \underbrace{(\innerp{M_1 - \beta_h}{x} + \epsilon)}_{\eqdef \eta_1(x)}. \label{eqn:y1}
\end{align}
The noise $\eta_1(x)$ is the sum of two terms:
(i) the \emph{mixing noise} $\innerp{M_1 - \beta_h}{x}$
due to the random choice of the mixture component $h$,
and (ii) the \emph{observation noise} $\epsilon \sim \normal{0}{\sigma^2}$.
Although the noise depends on $x$,
it still has zero mean conditioned on $x$.
We will later show that we can
perform linear regression on the data $\{\xni,
\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$
under identifiability conditions.
But clearly, knowing $M_1$ is insufficient
for identifying all the parameters $\theta$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theoretical results}

In this section, we provide theoretical guarantees for our Spectral Experts algorithm.
In particular, the following theorem provides a rate of convergence:

\begin{theorem}[Convergence of Spectral Experts]
\label{thm:convergence}
Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from a mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
Assume $d \ge k$.
Suppose the number of samples is
$n = \max(n_1,n_2)$
where $n_1 = \Omega\left(d^6 R^12 (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(d) \log(1/\delta)\right)$ and
$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
set to $\Omega(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}})$,
then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
\end{theorem}

The proof of the theorem has two parts.
First, we bound the error in the compound parameters estimates $\hat M_2,\hat M_3$
using results from \citet{Tomioka2011}.
Then, we use results from \cite{AnandkumarGeHsu2012} to convert this error
into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
derived from the robust tensor power method.
But first, let us study a more basic property: identifiability.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Identifiability from moments}

In ordinary linear regression, the regression coefficients $\beta \in \Re^d$ are
identifiable if and only if the data has full
rank: $\E[x\tp{2}] \succ 0$,
and furthermore, identifying $\beta$ requires only moments
$\E[xy]$ and $\E[x\tp{2}]$ (by observing the optimality conditions for \refeqn{y1}).
However, in mixture of linear regressions, these two moments only allow us to recover $M_1$.
\refthm{convergence} shows that if we have the higher order analogues,
$\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in \{1,2,3\}$,
we can then identify the parameters $\theta = (\pi, B)$,
provided $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for $p \in \{1,2,3\}$.

We note that the positive definite constraints are more stringent.
For example, suppose $x = (1, t, t^2)$,
the common polynomial basis expansion, so that all the coordinates are deterministically related.
Then $\E[x\tp{2}] \succ 0$ can easily be satisfied, say with $x \sim \normal{0}{I}$.
However, one can verify that $\E[\cvec(x\tp{2})\tp{2}]$ is singular,
since $\cvec(x\tp{2}) = [1 \cdot 1, t\cdot t, 2(1 \cdot t^2), 2(t \cdot t^2), (t^2 \cdot t^2)]$ contains
components $t \cdot t$ and $2(1 \cdot t^2)$, which are linearly dependent.
Therefore, Spectral Experts would not be able to identify the parameters of a mixture of
linear regressions for this distribution over $x$.

We now show that some amount of unidentifiability is intrinsic to estimation from low-order moments,
not just an artifact of our estimation procedure.
Suppose $x = (t, \dots, t^d)$.
Even if we observed all moments
$\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in [r]$,
all the resulting coordinates would be monomials up to degree $2dr$,
and thus the moments live in a $2(dr)$-dimensional subspace.
On the other hand, the parameters $\theta$ live in a subspace of at least dimension $dk$.
Therefore, at least $r \ge k/2$ moments are required for identifiability of any
algorithm for this monomial example.

>>>>

%%% first moment
To warm up, let us consider linear regression
on the response $y$ given $x$.
From the model definition, we have $y = \beta_h^\top x + \epsilon$ where
$\epsilon \sim \normal{0}{\sigma^2}$, and $h$ is a random quantity
independent of $x$.
We can average over this randomness by defining
the average regression coefficients
$M_1 \eqdef \sum_{h=1}^k \pi_h \beta_h$.
Now we can express $y$ as a linear function of $x$ with coefficients $M_1$
plus some noise $\eta_1(x)$:
\begin{align}
  y &= \innerp{M_1}{x} +
  \underbrace{(\innerp{\beta_h - M_1}{x} + \epsilon)}_{\eqdef \eta_1(x)}. \label{eqn:y1}
\end{align}
The noise $\eta_1(x)$ is the sum of two terms:
(i) the \emph{mixing noise} $\innerp{M_1 - \beta_h}{x}$
due to the random choice of the mixture component $h$,
and (ii) the \emph{observation noise} $\epsilon \sim \normal{0}{\sigma^2}$.
Although the noise depends on $x$,
it still has zero mean conditioned on $x$.
We will later show that we can
perform linear regression on the data $\{\xni,
\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$
under identifiability conditions.
But clearly, knowing $M_1$ is insufficient
for identifying all the parameters $\theta$.

@@

### 2013-04-23 doc/icml2013/evaluation.tex
Several typo/math fixes for the paper
Couple of corrections to the math in the theory and introduction section. Included a line about mode-k tensor unfoldings. Added details about $\sigma^2$ to the experiments

We generated synthetic data as follows:
First, we generated a vector $t$ sampled uniformly over the $b$-dimensional
unit hypercube $[-1,1]^b$.
Then, to get the actual covariates $x$, we applied a non-linear function of $t$
that conformed to the identifiability criteria discussed in
\sectionref{sec:algo}.
The coefficients $\{\beta_h\}$ were drawn from a standard Gaussian.

@@

\begin{table*}[p]
\caption{Parameter error $\|\theta^* - \hat \theta\|_F$ ($n = 500,000$)
as the number of variables $b$, number of features $d$ and the number of components $k$ increases.
While Spectral by itself does not produce good parameter estimates, Spectral+EM improves
over EM significantly.
}
\label{tbl:parameter-recovery}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}

@@

\begin{figure}[th]
  \centering
  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
  \label{fig:hist}
\end{figure}

@@

\begin{figure*}[p]
  \centering
  \subfigure[Well-specified data]{
    \includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.pdf}
  }
    \hspace{-2em}
  \subfigure[Misspecified data]{
    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
  }
  \caption{Learning curves: parameter error as a function of the number of samples $n$ ($b = 1, d = 5, k = 3$).}
  \label{fig:vs-n}
\end{figure*}

@@

\paragraph{Misspecified Data}

\begin{table*}[t]
\caption{Parameter rrror $\|\theta^* - \hat \theta\|_F$ when the data is misspecified ($n = 500,000$).
Spectral+EM degrades slightly, but still outperforms EM overall.
}
\label{tbl:parameter-recovery-mis}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}

>>>>

We generated synthetic data as follows:
First, we generated a vector $t$ sampled uniformly over the $b$-dimensional
unit hypercube $[-1,1]^b$.
Then, to get the actual covariates $x$, we applied a non-linear function of $t$
that conformed to the identifiability criteria discussed in
\sectionref{sec:algo}.
The coefficients $\{\beta_h\}$ were drawn from a standard Gaussian. The
results presented below use $\sigma^2 = 0.1$; we did not observe any
qualitatively different behaviour for choices of $\sigma^2$ in the range
$[0.01, 0.4]$.

@@

\begin{table*}[tbhp]
\caption{Parameter error $\|\theta^* - \hat \theta\|_F$ ($n = 500,000$)
as the number of variables $b$, number of features $d$ and the number of components $k$ increases.
While Spectral by itself does not produce good parameter estimates, Spectral+EM improves
over EM significantly.
}
\label{tbl:parameter-recovery}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}

@@

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
  \label{fig:hist}
\end{figure}

@@

\begin{figure*}[tbhp]
  \centering
  \subfigure[Well-specified data]{
    \includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.pdf}
  }
    \hspace{-2em}
  \subfigure[Misspecified data]{
    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
  }
  \caption{Learning curves: parameter error as a function of the number of samples $n$ ($b = 1, d = 5, k = 3$).}
  \label{fig:vs-n}
\end{figure*}

@@

\paragraph{Misspecified data}

\begin{table*}[tbhp]
\caption{Parameter error $\|\theta^* - \hat \theta\|_F$ when the data is misspecified ($n = 500,000$).
Spectral+EM degrades slightly, but still outperforms EM overall.
}
\label{tbl:parameter-recovery-mis}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}

### 2013-04-23 doc/icml2013/intro.tex
Couple of corrections to the math in the theory and introduction section. Included a line about mode-k tensor unfoldings. Added details about $\sigma^2$ to the experiments

We use $x\tp{p}$ to represent the $p$-th order tensor
formed by taking the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots
i_p} = x_{i_1} \cdots x_{i_p}$. We will use
$\innerp{\cdot}{\cdot}$ to denote the generalized dot product between two $p$-th
order tensors: $\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1,
\ldots i_p} Y_{i_1, \ldots i_p}.$
A tensor $X$ is symmetric if for all $i,j \in [d]^p$ which are permutations of each other,
$X_{i_1 \cdots i_p}$ = $X_{j_1 \cdots j_p}$ (all tensors in this paper will be symmetric).
For a $p$-th order tensor $X \in (\Re^d)\tp{p}$, let $X_{(i)} \in \Re^{d \times d^{p-1}}$
be the mode-$i$ unfolding of $X$.
%For $X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
%$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q}
%= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
%\ldots, i_q}$.

% Norms
For a matrix $X$,
let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
let $\|X\|_\text{op}$ denote the operator norm (largest singular value).
For a rank-$K$ tensor $X$,
let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
the average nuclear norm over all $p$ unfoldings,
and let $\|X\|_\text{op} = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\text{op}$
denote the average operator norm over all $p$ unfoldings.
Let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.

For a tensor $X \in (\Re^d)\tp{p}$,
let $\cvec(X) \in \Re^{\sum_{i=1}^p \binom{n}{i}}$ be
the collapsed vectorization of $X$: each component of $\cvec(X)$
is indexed by a vector of counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.
The value of that component is
$\sum_{k \in K(c)} X_{k_1 \cdots k_p}$,
where $K(c) = \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}| \}$ are the set of index vectors $k$ with that count profile.
For example, for $X \in \Re^{d \times d}$,
$\cvec(X) = (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

>>>>

We use $x\tp{p}$ to represent the $p$-th order tensor formed by taking
the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots i_p}
= x_{i_1} \cdots x_{i_p}$. We will use $\innerp{\cdot}{\cdot}$ to denote
the generalized dot product between two $p$-th order tensors:
$\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1, \ldots i_p} Y_{i_1,
\ldots i_p}$.  A tensor $X$ is symmetric if for all $i,j \in [d]^p$
which are permutations of each other, $X_{i_1 \cdots i_p}$ = $X_{j_1
\cdots j_p}$ (all tensors in this paper will be symmetric).  For
a $p$-th order tensor $X \in (\Re^d)\tp{p}$, the mode-$i$ unfolding of
$X$ is a matrix whose $j$-th row contains all the elements of $X$ whose
$i$-th index is equal to $j$. Let $X_{(i)} \in \Re^{d \times d^{p-1}}$
be the mode-$i$ unfolding of $X$.
%For $X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
%$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q} = \sum_{j_1,
%\ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1, \ldots,
%i_q}$.

% Norms
For a matrix $X$,
let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
let $\|X\|_\text{op}$ denote the operator norm (largest singular value).
For a tensor $X$,
let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
the average nuclear norm over all $p$ unfoldings,
and let $\|X\|_\text{op} = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\text{op}$
denote the average operator norm over all $p$ unfoldings.
Let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.

For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
\Re^{\binom{d + p + 1}{p}}$ be the collapsed vectorization of distinct
elements in $X$: each component of $\cvec(X)$ is indexed by a vector of
counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.  The value
of that component is $\sum_{k \in K(c)} X_{k_1 \cdots k_p}$, where $K(c)
= \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
\}$ are the set of index vectors $k$ with that count profile.  For
example, for $X \in \Re^{d \times d}$, $\cvec(X) = (X_{ii} : i \in [d];
X_{ij} + X_{ji} : i,j \in [d], i<j)$.
%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.
